---
output:
  word_document: default
  html_document: default
---
# Clustering
## Blaine Gibson

```{r}
library(tidyverse)
library(tidymodels)
library(skimr)
library(readr)

trucks <- read_csv("trucks.csv")
```
### Task 1
```{r}
ggplot(trucks,aes( x=Distance, y=Speeding)) +
  geom_point()
```

There appears to be around 4 natural clusters in each quadrant of the chart. The cluster in the bottom left quandrant appears to be the most dense.

### Task 2
```{r}
kmeans_recipe = recipe(~ Distance + Speeding, trucks) 

trucks1 = kmeans_recipe %>% 
  step_scale(all_numeric()) %>%
  step_center(all_numeric()) 

trucks1 = prep(trucks1, trucks) #prepares the recipe

trucks_cleaned = bake(trucks1, trucks) #applies the recipe and yields a data frame
```

```{r}
summary(trucks_cleaned)
summary(trucks)
```

### Task 3
```{r}
set.seed(64)
clusts = 
  tibble(k = 2) %>%
  mutate(
    kclust = map(k, ~kmeans(trucks_cleaned, .x)),
    tidied = map(kclust, tidy),
    glanced = map(kclust, glance),
    augmented = map(kclust, augment, trucks_cleaned)
  )

clusters = 
  clusts %>%
  unnest(cols = c(tidied))

assignments = 
  clusts %>% 
  unnest(cols = c(augmented))

clusterings = 
  clusts %>%
  unnest(cols = c(glanced))

p1 = 
  ggplot(assignments, aes(x = Distance, y = Speeding)) +
  geom_point(aes(color = .cluster), alpha = 0.8) + 
  facet_wrap(~ k)
p1
```
The clusters defined by k=2 show two clusters split among the distance variable but appear that they could be further clustered. The red cluster has two dominant groups and the blue cluster has a well defined lower cluster and a more varying higher speeding cluster.

### Task 4
```{r}
set.seed(412)
clusts2 = 
  tibble(k = 1:8) %>%
  mutate(
    kclust = map(k, ~kmeans(trucks_cleaned, .x)),
    tidied = map(kclust, tidy),
    glanced = map(kclust, glance),
    augmented = map(kclust, augment, trucks_cleaned)
  )
clusters2 = 
  clusts2 %>%
  unnest(cols = c(tidied))

assignments2 = 
  clusts2 %>% 
  unnest(cols = c(augmented))

clusterings2 = 
  clusts2 %>%
  unnest(cols = c(glanced))
p2 = 
  ggplot(assignments2, aes(x = Distance, y = Speeding)) +
  geom_point(aes(color = .cluster), alpha = 0.8) + 
  facet_wrap(~ k)
p2

```
The 4 cluster graph appears to be the best visual representation of the different clusters.

### Task 5
```{r}
ggplot(clusterings2, aes(k, tot.withinss)) +
  geom_line() +
  geom_point() + theme_bw()
```
A K value of 4 appears to be the optimal "elbow" for the amount of clusters. 

### Task 6
```{r}
clusts3 = 
  tibble(k = 4) %>%
  mutate(
    kclust = map(k, ~kmeans(trucks_cleaned, .x)),
    tidied = map(kclust, tidy),
    glanced = map(kclust, glance),
    augmented = map(kclust, augment, trucks_cleaned)
  )

clusters3 = 
  clusts %>%
  unnest(cols = c(tidied))

assignments3 = 
  clusts3 %>% 
  unnest(cols = c(augmented))

clusterings3 = 
  clusts3 %>%
  unnest(cols = c(glanced))

p3 = 
  ggplot(assignments3, aes(x = Distance, y = Speeding)) +
  geom_point(aes(color = .cluster), alpha = 0.8) + 
  facet_wrap(~ k)
p3
```

