---
output:
  word_document: default
  html_document: default
---
# Missing Data
## Blaine Gibson

```{r}
library(tidyverse)
library(VIM)
library(mice)
library(skimr)

grades <- read_csv("class-grades.csv")
str(grades)
summary(grades)
skim(grades)
```
Tutorial, Midterm, Take Home and Final columns are all missing data.

```{r}
vim_plot = aggr(grades, numbers = TRUE, prop = c(TRUE, FALSE),cex.axis=.7)

```

There are 10 students missing at least 1 peice of data including 1 student who is missing both the take home and midterm.  

Row-wise deletion:
```{r}
grades_rowdel = grades %>% drop_na()
str(grades_rowdel)
summary(grades_rowdel)
skim(grades_rowdel)
```

89 observation rows remain after row-wise deletion is applied.  

Column-wise deletion:
```{r}
grades_coldel = grades %>%  select(-Tutorial,-Midterm,-TakeHome,-Final)
vim_plot2 = aggr(grades_coldel, numbers = TRUE, prop = c(TRUE, FALSE),cex.axis=.7)
str(grades_coldel)
summary(grades_coldel)
skim(grades_coldel)
```
99 observations remain but with only 2 columns.  

In this instance, row-wise deletion seems to be the more logical option as it does not remove 4 variables from the data as column-wise deletion does. Column-wise deletion leaves very little of the data remaining to be analyzed.  
```{r}
set.seed(123)

imp_grades = mice(grades, m=5, method='pmm', printFlag=FALSE)

summary(imp_grades)

```

```{r}
grades_complete = complete(imp_grades)
summary(grades_complete)
```
The mean of the final variable has increased to 68.33 from 68.23, a rather minor change.
